### 2020-09-21 code_blocks/code_blocks.tex
Figure with code blocks

\documentclass{article}
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}}%
\AfterEndEnvironment{minted}{\end{tcolorbox}}%
\usepackage{url}

\usepackage[
    paperwidth=4.5in, % 27 pica for two column width in AJHG
    paperheight=7.5in,
    left=0.15in,
    right=0.075in,
    top=0.075in,
    bottom=0.05in
]{geometry}

\begin{document}

\begin{minted}[fontsize=\footnotesize]{tcsh}
-n 1 1.682020 -n 2 3.736830 -n 3 7.292050
-eg 0 2 116.010723 -eg 0 3 160.246047
-ma x 0.881098 0.561966 0.881098 x 2.797460 0.561966 2.797460 x
-ej 0.028985 3 2 -en 0.028985 2 0.287184
-ema 0.028985 3 x 7.293140 x 7.293140 x x x x x
-ej 0.197963 2 1 -en 0.303501 1 1
\end{minted}

\begin{minted}[linenos,highlightlines={16,17},numbersep=5pt,fontsize=\footnotesize]{python}
demographic_events = [
    # CEU and CHB merge into B with rate changes at T_EU_AS
    msprime.MassMigration(
        time=T_EU_AS, source=2, destination=1, proportion=1.0),
    msprime.MigrationRateChange(time=T_EU_AS, rate=0),
    msprime.MigrationRateChange(
        time=T_EU_AS, rate=m_AF_B, matrix_index=(0, 1)),
    msprime.MigrationRateChange(
        time=T_EU_AS, rate=m_AF_B, matrix_index=(1, 0)),
    msprime.PopulationParametersChange(
        time=T_EU_AS, initial_size=N_B, growth_rate=0,
        population_id=1),
    # Population B merges into YRI at T_B
    msprime.MassMigration(
        time=T_B, source=1, destination=0, proportion=1.0),
    # Missing in msprime documentation
    msprime.MigrationRateChange(time=T_B, rate=0),
    # Size changes to N_A at T_AF
    msprime.PopulationParametersChange(
        time=T_AF, initial_size=N_A, population_id=0)
]
\end{minted}

\begin{minted}[highlightlines={9},fontsize=\footnotesize]{python}
dbg = msprime.DemographyDebugger(
  population_configurations=population_configurations,
  demographic_events=demographic_events,
  migration_matrix=migration_matrix)
dbg.print_history()
ts = msprime.simulate(
  ..., # parameters defining samples, sequence length, etc
  population_configurations=population_configurations,
  demographic_events=demographic_events, # Missing in Ref. 11
  migration_matrix=migration_matrix)
\end{minted}

\begin{minted}[fontsize=\footnotesize]{python}
demography = msprime.Demography(
  populations=populations,
  events=demographic_events,
  migration_matrix=migration_matrix)
dbg = demography.debug()
dbg.print_history()
ts = msprime.simulate(
  ..., # parameters defining samples, sequence length, etc
  demography=demography)
\end{minted}

\end{document}